Article Title

### 2014-01-24 plos.tex
Kreso's edits of rev 41

\section*{Abstract}
% Please keep the abstract between 250 and 300 words
Ambitious projects aim to record the activity of ever larger and denser neuronal populations \emph{in vivo}.  Correlations in neural activity measured in such recordings can reveal important aspects of the organization of neural circuits.  However, estimating and interpreting large correlation matrices is statistically challenging.  Estimation can be improved by regularization, \ie by imposing a structure on the estimate.  The amount of improvement depends on how closely the assumed structure represents dependencies in the data. Therefore, the selection of the most efficient correlation matrix estimator for a given neural circuit must be determined empirically.  Importantly, the identity and structure of the most efficient estimator informs about the types of dominant dependencies governing the system.

We sought statistically efficient estimators of neural correlation matrices in recordings from large, dense groups of cortical neurons.  Using fast 3D random-access laser scanning microscopy of calcium signals, we recorded the activity of nearly every neuron in volumes of about 200 $\mu$m wide and 100 $\mu$m deep (150--350 cells) in mouse visual cortex.  We hypothesized that in these dense recordings, the correlation matrix should be best represented as the combination of a sparse matrix of pairwise partial correlations representing local interactions and a low-rank component representing common fluctuations and external inputs.  Indeed, in cross-validation tests, the covariance matrix estimator with this structure consistently outperformed other regularized estimators. The connectivity patterns inferred from the sparse component of the estimate were strongly related to the geometrical arrangement and orientation tuning properties of cells: the density of positive \sq{excitatory} connections decreased rapidly with geometric distances and with  differences in orientation preference whereas the negative \sq{inhibitory} connections were less selective.  Because of its superior performance, this estimator likely provides a more physiologically relevant representation of the functional connectivity in dense recordings than the sample correlation matrix.

@@

It is now possible to record the spiking activity of hundreds of neurons at the same time.  A meaningful statistical description of the collective activity of these neural populations -- their \sq{functional connectivity} -- is a forefront challenge in neuroscience.  We addressed this problem by identifying statistically efficient estimators of correlation matrices of the spiking activity of neural populations.  Various underlying processes may reflect differently on the structure of the correlation matrix:  Correlations due to common network fluctuations or external inputs are well estimated by low-rank representations, whereas correlations due to linear interactions between specific pairs of neurons are well approximated by their pairwise \emph{partial} correlations.  In our data obtained from fast 3D two-photon imaging of calcium signals of large and dense groups of neurons in mouse visual cortex, the best estimation performance was attained by decomposing the correlation matrix into a sparse network of partial correlations (\sq{interactions}) combined with a low-rank component. The inferred interactions were both positive (\sq{excitatory}) and negative (\sq{inhibitory}) and reflected the spatial organization and orientation preferences of the interacting cells.  We propose that the composition of the most efficient among many estimators provides a more informative picture of the functional connectivity than previous analyses of neural correlations.

\section*{Introduction}
\emph{Functional connectivity} is a statistical description of observed \emph{multineuronal} activity patterns not reducible to the response properties of the individual cells. Functional connectivity is thought to reflect local synaptic connections, shared inputs from other regions, and endogenous network activity.  Although functional connectivity is only a phenomenological description with no simple mechanistic interpretation, it can be used to generate or test hypotheses about the anatomical architecture of the neural circuit and the processing of information at the population level.

Pearson correlations between the spiking activity of pairs of neurons are among the most familiar measures of functional connectivity \cite{Averbeck:2006, Zohary:1994, Kohn:2005, Bair:2001, Ecker:2010}.  In particular, \emph{noise correlations}, \ie the correlations of trial-to-trial response variability between pairs of neurons, have a profound impact on stimulus coding \cite{Zohary:1994, Abbott:1999, Sompolinsky:2001, Nirenberg:2003, Averbeck:2006, Josic:2009, Berens:2011, Ecker:2011}. In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect aspects of the synaptic connectivity \cite{Gerstein:1964}.  The interest in neural correlations has been sustained by a series of discoveries of their nontrivial relationships to various other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008, Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ecker:2010, Rothschild:2010, Ko:2011, Smith:2013b}, cell types \cite{Hofer:2011}, cortical layer specificity \cite{Hansen:2012, Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009, Gu:2011, Ko:2013}, changes due to sensory stimulation and global brain states \cite{Greenberg:2008, Goard:2009, Kohn:2009, Rothschild:2010, Ecker:2010, Renart:2010} among others.

Neural correlations do not come with ready or unambiguous mechanistic interpretations. They can arise from monosynaptic or polysynaptic interactions, common or correlated inputs, oscillations, top-down modulation, and background network fluctuations \cite{Perkel:1967, Moore:1970, Shadlen:1998, Salinas:2001, Ostojic:2009, Rosenbaum:2011}. But multineuronal recordings do provide more information than available from an equivalent number of pairs of cells considered separately. For example, the eigenvalue decomposition of the covariance matrix expresses shared correlated activity components across the population; common fluctuations of population activity may be accurately represented by just a few principal components but will affect all correlation coefficients. On the other hand, a correlation matrix can be specified through the  partial correlations between pairs of the recorded neurons. The partial correlation coefficient between two neurons reflects their linear association conditioned on the activity of all the other recorded cells.  For this reason, partial correlations have been proposed to more closely reflect the direct causal effects between components of complex systems than correlations; partial correlations have been used to describe interactions between genes in functional genomics \cite{Schafer:2005, Peng:2009} and between brain regions in imaging studies \cite{Varoquaux:2012, Ryali:2012}. These opportunities have not yet been explored in neurophysiological studies: most studies have only consider the distribution of pairwise correlations \cite{Zohary:1994, Bair:2001, Smith:2008, Ecker:2010}.

However, estimation of covariance matrices from large populations presents a number of numerical challenges. The amount of recorded data grows only linearly with population size whereas the number of estimated coefficients increases quadratically. This mismatch leads to an increase in spurious correlations, overestimation of common activity (\ie overestimation of large eigenvalues) \cite{Ledoit:2004}, and poorly conditioned partial correlations \cite{Schafer:2005}. The \emph{sample correlation matrix} is an unbiased estimate of the true correlations but its many free parameters make it sensitive to the sampling noise. As a result, on average, the sample correlation matrix is farther from the true correlation matrix than some structured estimates.

Estimation can be improved through \emph{regularization},  the technique of deliberately imposing a structure on an estimate in order to reduce its estimation error \cite{Schafer:2005, Bickel:2006}. To \sq{impose a structure} on an estimate means to bias (\sq{shrink}) it toward a reduced representation with fewer free parameters, which is often referred to as the \emph{target estimate}.   The selection of the optimal target estimate and the optimal amount of shrinkage is deduced from the data sample either analytically \cite{Ledoit:2003, Ledoit:2004, Schafer:2005}  or by cross-validation \cite{Friedman:1989}. An estimator that produces estimates that are, on average, closer to the truth for a given sample size is said to be more \emph{efficient} than other estimators.

Although regularized covariance matrix estimation is commonplace in other fields such as finance \cite{Ledoit:2003}, functional genomics \cite{Schafer:2005}, and brain imaging \cite{Ryali:2012}, surprisingly little work has been done to identity optimal regularization of neural correlation matrices.

Improved estimation of the correlation matrix is beneficial in itself. For example, improved estimates can be used to optimize  decoding of the population activity. But reduced estimation error is not the only benefit of regularization.  The amount of improvement of estimation is greatest when the structure imposed by regularization limits the solution space but still represents the dominant dependencies in the data. Therefore, by finding the most efficient among many regularized estimators, we learn something about the system itself: the imposed structure of the most efficient estimator is a parsimonious representation of the regularities in the data. The elements of this structure likely provide a better representation of the underlying dependencies in the data than previous analyses of sample correlations.

The necessity and benefits of regularization increase with the size of the recorded population. With the advent of the age of big neural data, the search for optimal regularization schemes will become increasingly relevant for all models of functional connectivity. Since optimal regularization schemes are specific to systems under investigation, the inference of functional connectivity in big neural data will entail the search for optimal regularization schemes that may involve combinations of heuristic rules and numerical techniques specially designed for each neural circuit.

@@

\begin{FPfigure}
    \begin{center}
        \includegraphics{./figures/Figure02.pdf}
    \end{center}
    \caption{{\bf Regularized estimators whose structure matches the true structure in the data are more efficient.}
        {\bf Row 1.} Graphical representations of the target estimates used of the four respective regularized covariance matrix estimators.  Recorded neurons are denotes as green spheres and latent units as light-shaded spheres.  The edges between them denote non-zero partial correlations.
        {\bf Row 1, A}.  In estimator $C_{\sf diag}$, the target estimate is a diagonal matrix, which describes systems that lack linear associations. 
        {\bf  Row 1, B.} In estimator $C_{\sf factor}$, the target estimate is a factor model (low-rank matrix plus a diagonal matrix), favoring systems in which correlations arise due to common input from latent units. 
        {\bf  Row 1, C}. In estimator $C_{\sf sparse}$, the covariance matrix is approximated as the inverse of a sparse matrix. This approximation describes systems in which correlations arise from linear associations between the observed units.
        {\bf  Row 1, D}.  In estimator $C_{\sf sparse+latent}$, the covariance matrix is approximated as the inverse of the sum of a sparse matrix and a low-rank matrix. This approximation describes a model wherein correlations arise due to linear associations between the recorded cells \emph{and} with several latent units.
\\
{\bf Row 2:} Examples of $50\times 50$ correlation matrices corresponding to each structure: {\bf A.} the diagonal correlation matrix, {\bf B.} a factor model with two latent units, {\bf C.}  a correlation matrix with 85\%  off-diagonal zeros, and {\bf  D.} a correlation matrix whose inverse is the sum of a rank-1 (\ie one latent unit) matrix and a sparse matrix with 89\% off-diagonal zeros. 
\\
{\bf Row 3:} Sample correlation matrices calculated from samples of size $n=1000$ drawn from simulated random processes with respective correlation matrices from row 2.  The structure of the sample correlation matrix is difficult to discern.
\\
{\bf Row 4:} Correlation matrix estimates computed by estimators with matching structure from the same data as the sample correlation matrices in Row 3. The regularized estimates are closer to the truth than the sample correlation matrices.
\\
{\bf Row 5:} Excess losses (Eq.~\ref{eq:excess-loss}) attained by the five estimators as a function of sample size. The error bars indicate the standard error of the mean based on 30 samples.  Estimators with the matching structure converged to zero faster than the other estimators. Some exceptions occur when the sample is not large enough to reveal the true structure; in these cases estimators with simpler structures can be more efficient.
\\
{\bf Row 6:} Relative cross-validation losses (Eq.~\ref{eq:rel-cv-loss}) attained the five estimators with respect to the matching estimator. Error bars indicate the standard error of the mean based on 30 samples.
    }
    \label{fig:02}
\end{FPfigure}

The target estimate of estimator $C_{\sf diag}$ is the diagonal matrix $D$ containing estimates of neurons' variances. Regularization is achieved by linear \emph{shrinkage} of the unbiased estimate $C_{\sf 0}$ toward $D$ controlled by the scalar \emph{shrinkage intensity} parameter $\lambda \in [0, 1]$:
\begin{equation}\label{eq:c-diag}
C_{\sf diag} = (1-\lambda) C_{\sf 0} + \lambda D
\end{equation}
The structure imposed by $C_{\sf diag}$ favors populations with no linear associations between the neurons (\figref{02}{Row 1, A}).  If sample correlations arise from spurious effects, $C_{\sf diag}$ is expected to be more efficient than other estimators.

@@

Estimator $C_{\sf sparse}$ is produced by approximating the sample covariance matrix by a matrix whose inverse matrix is sparse: 
\begin{equation}\label{eq:c-sparse}
C_{\sf sparse} = S^{-1},
\end{equation}
where $S$ is a sparse matrix, \ie one in which a large fraction of off-diagonal elements are set to zero.  The estimator has one hyperparameter that determines the sparsity (fraction of off-diagonal zeros) in $S$.  The structure imposed by $C_{\sf sparse}$ favors conditions in which neural correlations arise from direct linear effects (\sq{interactions}) between some pairs of neurons (\figref{02}{Row 1, C}).  We will sometimes refer to sparse partial interactions between neurons as simply \sq{interactions}.

Estimator $C_{\sf sparse+latent}$ is produced by approximating the sample covariance matrix by a matrix whose inverse is the sum of a sparse component and a low-rank component: 
\begin{equation}\label{eq:c-sl}
C_{\sf sparse+latent} = (S - L)^{-1},
\end{equation}
where, as above, $S$ is a sparse matrix and $L$ is a low rank matrix. The estimator has two hyperparameters: one to regulate the sparsity of $S$ and the other to regulate the rank of $L$. The structure imposed by $C_{\sf sparse+latent}$ favors conditions in which the activity of neurons is determined by linear effects between some observed pairs of neurons and linear effects from several latent units (\figref{02}{Row 1, D}) \cite{Chandrasekaran:2010,Ma:2013}.

\paragraph{Simulation}
To demonstrate how the selection between regularized estimators allows discerning the structure of correlations, we constructed four $50\times 50$ covariance matrices with the structures imposed by the four respective regularized estimators (\figref{02}{\,Row 2, A--D}).  See the Methods section for details of how these matrices were generated.  We used these covariance matrices as the ground truth in multivariate normal distributions with zero means and drew samples of various sizes. Even with relatively large sample sizes (\eg $n=1000$), the sample correlation matrices were contaminated by sampling noise and their underlying structures were difficult to discern (\figref{02}{\,Row 3}).

The evaluation of a covariance matrix estimator $C$ is performed with respect to a \emph{loss function} $\loss{C,\Sigma}$ to quantify the discrepancy between the estimate and the truth, $\Sigma$.  The loss function must attain its minimum when $C=\Sigma$.

@@

This loss function derives from the multivariate normal log-likelihood function $L(\Sigma|C_{\sf 0})$ as:
\begin{equation}
     \loss{X,Y} \equiv -\ln(2\pi) - \frac 2 p L(X|Y)
\end{equation}
The normalization by $p$ makes the values of the loss function comparable across different population sizes.

Although the two functions appear similar in form, they differ conceptually: The log-likelihood $L(\Sigma|C_{\sf 0})$ is a function of the parameter $\Sigma$ given the sample covariance matrix $C_{\sf 0}$ whereas the loss function $\loss{C,\Sigma}$ expresses the deviation of an arbitrary estimate $C$ from the truth $\Sigma$.

@@

The \emph{excess loss} function is produced by shifting the loss function to zero at its minimum:
\begin{equation}\label{eq:excess-loss}
    \eloss{C,\Sigma} = \loss{C,\Sigma}-\loss{\Sigma,\Sigma},
\end{equation}

We drew 30 independent samples of size $n=250$, 500, 1000, 2000, and 4000 from model and computed the excess loss resulting from the five estimators.  The hyperparameters of the regularized estimators were optimized by nested cross-validation using only the data in the sample.  All the regularized estimators produced better estimates (lower excess loss) than the sample covariance matrix.  However, estimators whose structure matched the true model typically outperformed the other estimators (\figref{02}{\,row 5}).  There were two exceptions: First, for some smaller samples, the data were insufficient to reveal the true correlation structure and the estimators with simpler structures, $C_{\sf diag}$ and $C_{\sf factor}$, often outperformed the estimators with the matching structure (\figref{02}{\,Row 5, C and D}).  Second, $C_{\sf sparse+latent}$ often performed equally well to $C_{\sf sparse}$ on the data from models without latent units because it reliably inferred the right number of latent units, zero.

Even though, with increasing sample sizes, all estimators converged to the truth (zero excess loss), the distinction between the estimators did not decrease.

In empirical data, the ground truth, $\Sigma$, is not accessible and excess loss cannot be computed directly. Instead, the loss function can be estimated solely from the data through \emph{validation}.  In validation, an independent \emph{testing sample} is set aside to compute an additional sample covariance estimate $C_{\sf 0}^\prime$ to validate the estimate $C$ computed from the  \emph{training sample}.  Then \emph{validation loss} $\loss{C,C_{\sf 0}^\prime}$ can substitute the loss function $\loss{C,\Sigma}$.

The normal loss (Eq.~\ref{eq:loss}) is particularly convenient because it is additive in its second argument in the sense that 
 \begin{equation*}\label{eq:additivity}
 \loss{C,z_1} + \loss{C,z_2} \equiv \loss{C,z_1+z_2}
 \end{equation*}

In this case, the validation loss is an unbiased estimate of the loss function:
\begin{equation*}
    \E[C_{\sf 0}^\prime]{\loss{C,C_{\sf 0}^\prime}}=\loss{C,\E[C_{\sf 0}^\prime]{C_{\sf 0}^\prime}}=\loss{C,\Sigma}
\end{equation*}

@@

Acquiring a separate testing sample is not practical. Instead, $K$-fold cross-validation can be used: The sample is divided into $K$ subsets of approximately equal size ($K=10$ in this study).  Then each subset is used as the validation sample with the other $K-1$ serving as the training sample. The validation losses from each of such \sq{folds} are averaged to produce the \emph{cross-validation loss} or \emph{CV-loss} for short.  Let $C_{\sf 0}^{\{k\}}$ denote the sample covariance matrix computed from the $k^{th}$ subset and $C^{\{\setminus k\}}$ denote the results of estimator $C$ trained on the remaining $K-1$ subsets. Then the cross-validation loss for estimator $C$ is
\begin{equation}\label{eq:cv-loss}
    \ell_C=\frac 1 K \sum\limits_{k=1}^K \loss{C^{\{\setminus k\}},C_{\sf 0}^{\{k\}}}
\end{equation}
In this formulation, $\ell_C$ can be recognized as the cross-validated Gaussian log-likelihood (up to a constant offset and multiplier). However, the present framework is more general, allowing for other loss functions to be used.

>>>>

\section*{Abstract}
% Please keep the abstract between 250 and 300 words
Ambitious projects aim to record the activity of ever larger and denser neuronal populations \emph{in vivo}.  Correlations in neural activity measured in such recordings can reveal important aspects of  neural circuit organization.  However, estimating and interpreting large correlation matrices is statistically challenging.  Estimation can be improved by regularization, \ie by imposing a structure on the estimate.  The amount of improvement depends on how closely the assumed structure represents dependencies in the data. Therefore, the selection of the most efficient correlation matrix estimator for a given neural circuit must be determined empirically.  Importantly, the identity and structure of the most efficient estimator informs about the types of dominant dependencies governing the system.

We sought statistically efficient estimators of neural correlation matrices in recordings from large, dense groups of cortical neurons.  Using fast 3D random-access laser scanning microscopy of calcium signals, we recorded the activity of nearly every neuron in volumes of about 200 $\mu$m wide and 100 $\mu$m deep (150--350 cells) in mouse visual cortex.  We hypothesized that in these dense recordings, the correlation matrix should be best represented as the combination of a sparse matrix of pairwise partial correlations representing local interactions and a low-rank component representing common fluctuations and external inputs.  Indeed, in cross-validation tests, the covariance matrix estimator with this structure consistently outperformed other regularized estimators. The connectivity patterns inferred from the sparse component of the estimate were strongly related to the geometrical arrangement and orientation tuning properties of the recorded cells: the density of positive \sq{excitatory} connections decreased rapidly with geometric distances and with  differences in orientation preference whereas the negative \sq{inhibitory} connections were less selective.  Because of its superior performance, this ``sparse + latent'' estimator likely provides a more physiologically relevant representation of the functional connectivity in dense recordings than the sample correlation matrix.

@@

It is now possible to record the spiking activity of hundreds of neurons at the same time.  A meaningful statistical description of the collective activity of these neural populations which reveals their \sq{functional connectivity} is a forefront challenge in neuroscience.  We addressed this problem by identifying statistically efficient estimators of correlation matrices of the spiking activity of neural populations.  Various underlying processes may reflect differently on the structure of the correlation matrix:  Correlations due to common network fluctuations or external inputs are well estimated by low-rank representations, whereas correlations due to linear interactions between specific pairs of neurons are well approximated by their pairwise \emph{partial} correlations.  In our data obtained from fast 3D two-photon imaging of calcium signals of large and dense groups of neurons in mouse visual cortex, the best estimation performance was attained by decomposing the correlation matrix into a sparse network of partial correlations (\sq{interactions}) combined with a low-rank component. The inferred interactions were both positive (\sq{excitatory}) and negative (\sq{inhibitory}) and reflected the spatial organization and orientation preferences of the interacting cells.  We propose that  the most efficient among many estimators provides a more informative picture of the functional connectivity than previous analyses of neural correlations.

\section*{Introduction}
\emph{Functional connectivity} is a statistical description of observed \emph{multineuronal} activity patterns not reducible to the response properties of the individual cells. Functional connectivity reflects local synaptic connections, shared inputs from other regions, and endogenous network activity. Although functional connectivity is a phenomenological description with no simple mechanistic interpretation, it can be used to generate hypotheses about the anatomical architecture of the neural circuit and test hypotheses about the processing of information at the population level.

Pearson correlations between the spiking activity of pairs of neurons are among the most familiar measures of functional connectivity \cite{Averbeck:2006, Zohary:1994, Kohn:2005, Bair:2001, Ecker:2010}.  In particular, \emph{noise correlations}, \ie the correlations of trial-to-trial response variability between pairs of neurons, have a profound impact on stimulus coding \cite{Zohary:1994, Abbott:1999, Sompolinsky:2001, Nirenberg:2003, Averbeck:2006, Josic:2009, Berens:2011, Ecker:2011}. In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect aspects of synaptic connectivity \cite{Gerstein:1964}.  Interest in neural correlations has been sustained by a series of discoveries of their nontrivial relationships to various aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008, Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ecker:2010, Rothschild:2010, Ko:2011, Smith:2013b}, cell types \cite{Hofer:2011}, cortical layer specificity \cite{Hansen:2012, Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009, Gu:2011, Ko:2013}, changes due to sensory stimulation and global brain states \cite{Greenberg:2008, Goard:2009, Kohn:2009, Rothschild:2010, Ecker:2010, Renart:2010}.

Neural correlations do not come with ready or unambiguous mechanistic interpretations. They can arise from monosynaptic or polysynaptic interactions, common or correlated inputs, oscillations, top-down modulation, and background network fluctuations, or a combination of these and other mechanisms \cite{Perkel:1967, Moore:1970, Shadlen:1998, Salinas:2001, Ostojic:2009, Rosenbaum:2011}. But multineuronal recordings do provide more information than an equivalent number of separately recorded pairs of cells. For example, the eigenvalue decomposition of the covariance matrix expresses shared correlated activity components across the population; common fluctuations of population activity may be accurately represented by just a few principal components but will affect all correlation coefficients. On the other hand, a correlation matrix can be specified using the emph{partial correlations} between pairs of the recorded neurons. The partial correlation coefficient between two neurons reflects their linear association conditioned on the activity of all other recorded cells.  For this reason, partial correlations more closely reflect the direct causal effects between components of complex systems than correlations; partial correlations have been used to describe interactions between genes in functional genomics \cite{Schafer:2005, Peng:2009} and between brain regions in imaging studies \cite{Varoquaux:2012, Ryali:2012}. These opportunities have not yet been explored in neurophysiological studies, as most studies have only consider the distribution of pairwise correlations \cite{Zohary:1994, Bair:2001, Smith:2008, Ecker:2010}.

However, estimation of covariance matrices from large populations presents a number of numerical challenges. The amount of recorded data grows only linearly with population size whereas the number of estimated coefficients increases quadratically. This mismatch leads to an increase in spurious correlations, overestimation of common activity (\ie overestimation of large eigenvalues) \cite{Ledoit:2004}, and poorly conditioned partial correlations \cite{Schafer:2005}. The \emph{sample correlation matrix} is an unbiased estimate of the true correlations but its many free parameters make it sensitive to sampling noise. As a result, on average, the sample correlation matrix is farther from the true correlation matrix than some structured estimates.

Estimation can be improved through \emph{regularization},  the technique of deliberately imposing a structure on an estimate in order to reduce its estimation error \cite{Schafer:2005, Bickel:2006}. To \sq{impose a structure} on an estimate means to bias (\sq{shrink}) it toward a reduced representation  with fewer free parameters, which is often referred to as the \emph{target estimate}.   The optimal target estimate and the optimal amount of shrinkage can be obtained from the data sample either analytically \cite{Ledoit:2003, Ledoit:2004, Schafer:2005},  or by cross-validation \cite{Friedman:1989}. An estimator that produces estimates that are, on average, closer to the truth for a given sample size is said to be more \emph{efficient} than other estimators.

Although regularized covariance matrix estimation is commonplace in finance \cite{Ledoit:2003}, functional genomics \cite{Schafer:2005}, and brain imaging \cite{Ryali:2012}, surprisingly little work has been done to identify optimal regularization of neural correlation matrices.

Improved estimation of the correlation matrix is beneficial in itself. For example, improved estimates can be used to optimize  decoding of the population activity. But reduced estimation error is not the only benefit of regularization.  The amount of improvement of estimation is greatest when the structure imposed by regularization limits the solution space, but still represents the dominant dependencies in the data. Therefore, by finding the most efficient among many regularized estimators, we learn something about the system itself: the structure of the most efficient estimator is a parsimonious representation of the regularities in the data.

The necessity and benefits of regularization increase with the size of the recorded population. With the advent of  big neural data, the search for optimal regularization schemes will become increasingly relevant for all models of functional connectivity. Since optimal regularization schemes are specific to systems under investigation, the inference of functional connectivity in large scale neural data will entail the search for optimal regularization schemes that may involve combinations of heuristic rules and numerical techniques specially designed for each neural circuit.

@@

\begin{FPfigure}
    \begin{center}
        \includegraphics{./figures/Figure02.pdf}
    \end{center}
    \caption{{\bf Regularized estimators whose structure matches the true structure in the data are more efficient.}
        {\bf Row 1.} Graphical representations of the target estimates of the four respective regularized covariance matrix estimators.  Recorded neurons are represented by green spheres and latent units by light-shaded spheres.  Edges represent non-zero partial correlations.
        {\bf Row 1, A}.  For estimator $C_{\sf diag}$, the target estimate is a diagonal matrix, which describes systems that lack linear associations. 
        {\bf  Row 1, B.} For estimator $C_{\sf factor}$, the target estimate is a factor model (low-rank matrix plus a diagonal matrix), representing systems in which correlations arise due to common input from latent units. 
        {\bf  Row 1, C}. For estimator $C_{\sf sparse}$, the covariance matrix is approximated as the inverse of a sparse matrix. This approximation describes systems in which correlations arise from a sparse set of  linear associations between the observed units. 
        {\bf  Row 1, D}.  For estimator $C_{\sf sparse+latent}$, the covariance matrix is approximated as the inverse of the sum of a sparse matrix and a low-rank matrix. This approximation describes a model wherein correlations arise due to sparse associations between the recorded cells \emph{and} with several latent units. \\
{\bf Row 2:} Examples of $50\times 50$ correlation matrices corresponding to each structure: {\bf A.} the diagonal correlation matrix, {\bf B.} a factor model with two latent units, {\bf C.}  a correlation matrix with 85\%  off-diagonal zeros, and {\bf  D.} a correlation matrix whose inverse is the sum of a rank-1 (\ie one latent unit) matrix and a sparse matrix with 89\% off-diagonal zeros. 
\\
{\bf Row 3:} Sample correlation matrices calculated from samples of size $n=1000$ drawn from simulated random processes with respective correlation matrices from row 2.  The structure of the sample correlation matrix is difficult to discern.
\\
{\bf Row 4:} Correlation matrix estimates computed by estimators with matching structure from the same data as the sample correlation matrices in Row 3. The regularized estimates are closer to the truth than the sample correlation matrices.
\\
{\bf Row 5:} Excess losses (Eq.~\ref{eq:excess-loss}) for the five estimators as a function of sample size. The error bars indicate the standard error of the mean based on 30 samples.  Estimators with structure that matches the model used to generate the data converged to zero faster than the other estimators. Some exceptions occur when the sample is too small to reveal the structure in the data. In these cases estimators with simpler structures can be more efficient.
\\
{\bf Row 6:} Relative cross-validation losses (Eq.~\ref{eq:rel-cv-loss}) for the five estimators with respect to the matching estimator. Error bars indicate the standard error of the mean based on 30 samples.
    }
    \label{fig:02}
\end{FPfigure}

The target estimate of estimator $C_{\sf diag}$ is the diagonal matrix $D$ containing estimates of neurons' variances. Regularization is achieved by linear \emph{shrinkage} of the unbiased estimate $C_{\sf 0}$ toward $D$ controlled by the scalar \emph{shrinkage intensity} parameter $\lambda \in [0, 1]$:
\begin{equation}\label{eq:c-diag}
C_{\sf diag} = (1-\lambda) C_{\sf 0} + \lambda D
\end{equation}
The structure imposed by $C_{\sf diag}$ favors (performs better  on) populations   with no linear associations between the neurons (\figref{02}{Row 1, A}).  If sample correlations arise from spurious effects, $C_{\sf diag}$ is expected to be more efficient than other estimators.

@@

Estimator $C_{\sf sparse}$ is produced by approximating the sample covariance matrix whose inverse is sparse: 
\begin{equation}\label{eq:c-sparse}
C_{\sf sparse} = S^{-1}.
\end{equation}
Here $S$ is a sparse matrix, \ie one in which a large fraction of off-diagonal elements are set to zero.  The estimator has one hyperparameter that determines the sparsity (fraction of off-diagonal zeros) in $S$.  The structure imposed by $C_{\sf sparse}$ favors conditions in which neural correlations arise from direct linear effects (\sq{interactions}) between some pairs of neurons (\figref{02}{Row 1, C}).  We will sometimes refer to sparse partial correlations between neurons simply  as \sq{interactions}.

Estimator $C_{\sf sparse+latent}$ is obtained by approximating the sample covariance matrix by a matrix whose inverse is the sum of a sparse component and a low-rank component: 
\begin{equation}\label{eq:c-sl}
C_{\sf sparse+latent} = (S - L)^{-1},
\end{equation}
where, as above, $S$ is a sparse matrix and $L$ is a low rank matrix. The estimator has two hyperparameters: one to regulate the sparsity of $S$ and the other to regulate the rank of $L$. The structure imposed by $C_{\sf sparse+latent}$ favors conditions in which the activity of neurons is determined by linear effects between some observed pairs of neurons and linear effects from several latent units (\figref{02}{Row 1, D}) \cite{Chandrasekaran:2010,Ma:2013}.

\paragraph{Simulation}
We next demonstrat how the most efficient among different regularized estimators can reveal the structure of correlations.
We constructed four $50\times 50$ covariance matrices, each with structure that matched one of the four regularized estimators (\figref{02}{\,Row 2, A--D}, see Methods for details about how these matrices were generated).  We used these covariance matrices as the ground truth in multivariate normal distributions with zero means and drew samples of various sizes. Even with relatively large sample sizes (\eg $n=1000$), the sample correlation matrices were contaminated by sampling noise and their underlying structures were difficult to discern (\figref{02}{\,Row 3}).

The evaluation of a covariance matrix estimator, $C,$ is performed with respect to a \emph{loss function} $\loss{C,\Sigma}$ to quantify the discrepancy between the estimate and the truth, $\Sigma$.  The loss function must attain its minimum when $C=\Sigma$.

@@

This loss function is related to the multivariate normal log-likelihood function $L(\Sigma|C_{\sf 0})$,
\begin{equation}
     \loss{X,Y} \equiv -\ln(2\pi) - \frac 2 p L(X|Y)
\end{equation}
Normalization by $p$ makes the values of the loss function comparable across different population sizes.

Although the two functions are similar in form, they differ conceptually: The log-likelihood $L(\Sigma|C_{\sf 0})$ is a function of the parameter $\Sigma$ given the sample covariance matrix $C_{\sf 0}$ whereas the loss function $\loss{C,\Sigma}$ expresses the deviation of an arbitrary estimate $C$ from the truth $\Sigma$.

@@

The \emph{excess loss} function is produced by shifting the loss function to zero at its minimum:
\begin{equation}\label{eq:excess-loss}
    \eloss{C,\Sigma} = \loss{C,\Sigma}-\loss{\Sigma,\Sigma}
\end{equation}

We drew 30 independent samples of size $n=250$, 500, 1000, 2000, and 4000 from model and computed the excess loss for each of the five estimators.  The hyperparameters of the regularized estimators were optimized by nested cross-validation using only the data in the sample.  All the regularized estimators produced better estimates (lower excess loss) than the sample covariance matrix.  However, estimators whose structure matched the true model typically outperformed the other estimators (\figref{02}{\,row 5}).  There were two exceptions: First, for some smaller samples, the data were insufficient to reveal the true correlation structure and the estimators with simpler structures, $C_{\sf diag}$ and $C_{\sf factor}$, often outperformed the estimators with the matching structure (\figref{02}{\,Row 5, C and D}).  Second, $C_{\sf sparse+latent}$ often performed as well as $C_{\sf sparse}$ on data from models without latent units because.  However, in theses cases it reliably inferred the correct number of latent units which was zero.

With increasing sample sizes, all estimators converged to the truth (zero excess loss).  However, 
the discriminability between their performances increased with sample size (data not shown).

With empirical data, the ground truth, $\Sigma$, is not accessible and excess loss cannot be computed directly. Instead, the loss function must be estimated from the sample through \emph{validation}.  In a validation procedure, an independent \emph{testing sample} is set aside to compute an additional sample covariance estimate $C_{\sf 0}^\prime$ to validate the estimate $C$ computed from the  \emph{training sample}.   The estimate $C_{\sf 0}^\prime$ is used to 
compute the \emph{validation loss} $\loss{C,C_{\sf 0}^\prime}$, which approximates the loss  $\loss{C,\Sigma}$.

The normal loss (Eq.~\ref{eq:loss}) is particularly convenient because it is additive in its second argument, \ie
 \begin{equation*}\label{eq:additivity}
 \loss{C,z_1} + \loss{C,z_2} \equiv \loss{C,z_1+z_2}
 \end{equation*}
With this property of additivity, the validation loss is an unbiased estimate of the loss function:
\begin{equation*}
    \E[C_{\sf 0}^\prime]{\loss{C,C_{\sf 0}^\prime}}=\loss{C,\E[C_{\sf 0}^\prime]{C_{\sf 0}^\prime}}=\loss{C,\Sigma}.
\end{equation*}

@@

Acquiring a separate testing sample is not practical. Instead, $K$-fold cross-validation can be used: The sample is divided into $K$ subsets of approximately equal size ($K=10$ in this study).  Then each subset is used as the validation sample with the other $K-1$ serving as the training sample. The validation losses from all such \sq{folds} are averaged to produce the \emph{cross-validation loss} or \emph{CV-loss} for short.  Let $C_{\sf 0}^{\{k\}}$ denote the sample covariance matrix computed from the $k^{th}$ subset and $C^{\{\setminus k\}}$ denote the results of estimator $C$ when the remaining $K-1$ subsets are treated as the training sample. Then the cross-validation loss for estimator $C$ is
\begin{equation}\label{eq:cv-loss}
    \ell_C=\frac 1 K \sum\limits_{k=1}^K \loss{C^{\{\setminus k\}},C_{\sf 0}^{\{k\}}}
\end{equation}
In this formulation, $\ell_C$ can be recognized as the cross-validated Gaussian log-likelihood (up to a constant offset and multiplier). However, the present framework is more general, as other loss functions can be used.

Since we only need to compare estimators to each other, we are  interested in the \emph{relative} CV loss of estimator $C$ with respect to reference estimator $C_{\sf ref}$,
\begin{equation}\label{eq:rel-cv-loss}
    \Delta\ell_{C,C_{\sf ref}} = \ell_{C}-\ell_{C_{\sf ref}}
\end{equation}
The relative cross-validation loss was used for the main findings in this study.
